Donor Preferences in Canadian Political Financing

### 2024-12-04 paper/paper.qmd
updated summary tables

# Data {#sec-data}

@@

\newpage
```{r tbl-fed-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-fed-donation
#| tbl-cap: Table Showing Donations Based on Recipient Power Status at Federal Level

@@

# Render Ontario Donations Table
kable(donations_in_power_federal, 
      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
      format = "latex", 
      booktabs = TRUE,  
      align = c("l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

@@

# Render Ontario Donations Table
kable(donations_in_power_ontario, 
      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
      format = "latex", # For PDF/LaTeX output
      booktabs = TRUE,  # Add LaTeX booktabs style
      align = c("l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

@@

```{r fig-ontario-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-ontario-total-donation
#| tbl-cap: Figure Showing Total Donations by Political Party at Ontario Level

@@

```{r fig-federal-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-federal-total-donation
#| tbl-cap: Figure Showing Total Donations by Political Party at Federal Level

@@

To analyze the impact of a political party’s power status on donation
patterns within Canadian federal and Ontario provincial elections,
multiple linear regression models were developed. Specifically, separate
models were constructed for Ontario and Federal levels to account for
regional differences in political dynamics and donation behaviors.
Additionally, a focused regression model was created for the
Conservative and Liberal parties to explore the interaction between
power status and party size.

@@

To analyze the impact of a political party's power status on donation
patterns, we employed linear regression models for both Ontario and
federal levels:

@@

Where:
- \(\text{Party}_i\): A binary variable indicating 1 for Conservative, 0 for Liberal.\\
```
Each model uses the logarithm of total donations (Log_Total_Donations)
as the dependent variable to stabilize variance and normalize the
distribution of donation amounts. Predictor variables include binary
indicators for whether the party was in power (In_Power), the size of
the party (Party_Size), and whether the donation was made during an
election year (Election_Year). The specialized model for Conservative
and Liberal parties includes an additional binary variable (Party) and
an interaction term (In_Power \* Party) to capture the combined effect
of power status and party type on donation amounts.

@@

The models assume a linear relationship between the predictors and the
logarithm of total donations, ensuring that the effects of the
predictors are additive and proportional. Additionally, they assume that
the residuals are normally distributed (normality of errors) and exhibit
constant variance (homoscedasticity). The models also rely on the
absence of multicollinearity, meaning that the predictors are not highly
correlated, and they assume independence of observations, where each
data point is independent of the others. Potential limitations of these
models include the possibility of unobserved variables influencing
donation amounts and the inherent assumption that past trends can
reliably predict future behavior.

### Model Validation \*\*\*\*\*\*

@@

For the Ontario Model, the regression analysis yielded an R² of 0.645
and an Adjusted R² of 0.635, indicating that approximately 64.5% of the
variance in log-transformed total donations is explained by the
predictors. The RMSE for this model was calculated to assess the average
prediction error, demonstrating a reasonable level of accuracy.

The Federal Model showed an R² of 0.515 and an Adjusted R² of 0.503,
suggesting that 51.5% of the variance in log-transformed total donations
is accounted for by the predictors. The RMSE value indicated acceptable
predictive performance, though slightly lower than the Ontario model,
reflecting potential differences in donation dynamics at the federal
level.

In the Conservative and Liberal Parties Model, the inclusion of the
interaction term between In_Power and Party resulted in an R² of 0.630
and an Adjusted R² of 0.600. This model effectively captures the
combined effect of power status and party type on donation amounts, with
the interaction term providing significant insights into how being in
power influences donations differently for Conservative and Liberal
parties. The RMSE for this model confirmed its robust predictive
capability, comparable to the other models.

Overall, the validation process confirmed that the linear regression
models are robust and effective in predicting donation amounts based on
the selected predictor variables. The models demonstrate good
explanatory power and reasonable predictive accuracy, making them
suitable for analyzing the factors influencing political donations in
Canada. However, ongoing evaluation and potential model refinement are
recommended to accommodate evolving political and financial landscapes,
ensuring the models remain relevant and accurate over time.

@@

Table @tbl-modelsummary presents the summary of key model estimates for
the Ontario, Federal, and Conservative/Liberal models. These models
predict the logarithm of total donations received by political parties
based on whether the party was in power, the size of the party, and
whether the donation occurred during an election year. The
Conservative/Liberal model includes an interaction term to explore how
the effect of being in power varies between these two major parties.

>>>>

\newpage
# Data {#sec-data}

@@

```{r tbl-fed-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-fed-donation
#| tbl-cap: Table Showing Donations Based on Recipient Power Status at Federal Level

@@

# Render Ontario Donations Table
kable(donations_in_power_federal, 
      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
      format = "latex", 
      booktabs = TRUE,  
      align = c("l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped"))

@@

# Render Ontario Donations Table
kable(donations_in_power_ontario, 
      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
      format = "latex", # For PDF/LaTeX output
      booktabs = TRUE,  # Add LaTeX booktabs style
      align = c("l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped"))
```

@@

```{r fig-ontario-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-ontario-total-donation
#| fig-cap: Figure Showing Total Donations by Political Party at Ontario Level

@@

```{r fig-federal-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-federal-total-donation
#| fig-cap: Figure Showing Total Donations by Political Party at Federal Level

@@

To analyze the impact of a political party’s power status on donation
patterns within Canadian federal and Ontario provincial elections,
multiple linear regression models were developed. Specifically, separate
models were constructed for Ontario and Federal levels to account for
regional differences in political dynamics and donation behaviors.
Additionally, a focused regression model was created for the
Conservative and Liberal parties to explore the interaction between power status and party affiliation.

@@

The models are specified as follows for the Ontario and Federal Models:

@@

Where:
- \(\text{Party}_i\): A binary variable indicating 1 for Conservative, 0 for Liberal.\\
```
Each model uses the logarithm of total donations (Log_Total_Donations)
as the dependent variable to stabilize variance and normalize the
distribution of donation amounts. Predictor variables include binary
indicators for whether the party was in power (In_Power), the size of
the party (Party_Size), and whether the donation was made during an
election year (Election_Year).

The specialized model for Conservative
and Liberal parties includes an additional binary variable (Party) and
an interaction term (In_Power \* Party) to capture the combined effect
of power status and party type on donation amounts.

@@

The models assume a linear relationship between the predictors and the logarithm of total donations, ensuring that the effects of the predictors are additive and proportional. Additionally, they assume that the residuals are normally distributed and exhibit constant variance (homoscedasticity). The models also rely on the absence of multicollinearity, meaning that the predictors are not highly correlated, and they assume independence of observations.

Potential limitations include the possibility of unobserved variables influencing donation amounts and the inherent assumption that past trends can reliably predict future behavior. These factors could introduce bias or limit the generalizability of the findings.

### Model Validation

@@

The Ontario model yielded an R² of 0.5496 and an Adjusted R² of 0.5267, indicating that approximately 54.96% of the variance in the log-transformed total donations is explained by the predictors. The RMSE was 2.5919, reflecting the average prediction error. The Variance Inflation Factor (VIF) values for all predictors were below 2, suggesting no significant multicollinearity.

The Federal model showed an R² of 0.6487 and an Adjusted R² of 0.6410, indicating that about 64.87% of the variance is explained by the model. The RMSE was 1.4909, and VIF values were again below 2, suggesting no significant multicollinearity.

The Conservative/Liberal model achieved an R² of 0.1091 and an Adjusted R² of 0.0431, indicating a lower explanatory power compared to the previous models. The RMSE was 1.2814. VIF values suggested moderate multicollinearity due to the interaction term, but they were within acceptable limits.

Overall, the validation process confirmed that the linear regression models are effective in predicting donation amounts based on the selected predictor variables. The models demonstrate good explanatory power and reasonable predictive accuracy, making them suitable for analyzing the factors influencing political donations in Canada.

@@

@tbl-modelsummary presents the summary of key model estimates for
the Ontario, Federal, and Conservative/Liberal models. These models
predict the logarithm of total donations received by political parties
based on whether the party was in power, the size of the party, and
whether the donation occurred during an election year. The
Conservative/Liberal model includes an interaction term to explore how
the effect of being in power varies between these two major parties.
